---
title: "Reactive Results of Chick Weight Experiment"
author: "Alicia Korol"
date: "March 30, 2018"
output: ioslides_presentation
smaller: true
theme: sky
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Chick weight experiment overview 

Experiment described in R:

`?ChickWeight`

<div class="red">
Description

The body weights of the chicks were measured at birth and every second day thereafter until day 20. They were also measured on day 21. There were four groups on chicks on different protein diets.
</div>


## Data set components 

Variables in the data frame include: 

- `ChickWeight$weight`
   
    body weights of chicks

- `ChickWeight$Time`
   
    day of experiment

- `ChickWeight$Chick`
    
    unique identifier given to each of 50 chicks

- `ChickWeight$Diet`
    
    ID of protein diet fed to chicks

## Challenge: Creating compact visuals

Drawbacks of static plots for complex (e.g., grouped) data

- Multiple redundant plots needed to display one data set
- Large space requirements in HTML documents
- Plot size is often reduced

<div class="blue">
**SOLUTION = reactive plots** 
</div>

```{r packages, echo = FALSE}
library(shiny)
library(dplyr)
library(tidyr)
```

----
```{r app}
data(ChickWeight)
diet1 <- ChickWeight %>%
    filter(Diet==1)
diet2 <- ChickWeight %>%
    filter(Diet==2)
diet3 <- ChickWeight %>%
    filter(Diet==3)
diet4 <- ChickWeight %>%
    filter(Diet==4) 

ui <- fluidPage(
    
    titlePanel("Effect of different protein diets on chick weights"),
    
    # Sidebar with a slider input for number of bins 
    sidebarLayout(
        sidebarPanel(
            selectInput(inputId = "DietSelect",label= "Select Diet",
                        choices=c("Diet 1","Diet 2","Diet 3","Diet 4")), #see selectInput?
            #Input name for internal use; label=displayed label/title
            submitButton("Submit")
        ),
        
        # Show a plot of the generated distribution
        mainPanel(
            plotOutput(outputId = "plot1"), #id to be used internally
            h3("Mean change in chick weight (g) after 21 days:"),
            textOutput("stats")
        )
    )
)

server <- function(input,output){
    
    output$plot1 <- renderPlot({
        
        data <- switch(input$DietSelect,
                       "Diet 1" = diet1,
                       "Diet 2" = diet2,
                       "Diet 3" = diet3,
                       "Diet 4" = diet4)
        
        # convert factor to numeric for convenience 
        data$Chick <- as.numeric(data$Chick)
        nChick <- length(unique(data$Chick))
        
        # get the range for the x and y axis 
        xrange <- range(data$Time)
        yrange <- c(25,400)
        
        #Set up plot
        plot(xrange, yrange, type="n", xlab="Time (days)",
             ylab="Weight (g)" ) 
        colors <- rainbow(nChick) 
        linetype <- rep(1,nChick) 
        plotchar <- rep(18,nChick)
        
        # add lines 
        if(sort(unique(data$Chick))[1]=="1"){m <- 0
        } else if (sort(unique(data$Chick))[1]=="21") {m <- 20
        } else if (sort(unique(data$Chick))[1]=="31") {m <- 30
        } else m <- 41
        
        for (i in 1:nChick) { 
            sub <- subset(data, Chick==i+m) 
            lines(sub$Time, sub$weight, type="b", lwd=1.5,
                  lty=linetype[i], col=colors[i], pch=plotchar[i]) 
        } 
        
        # add a title and subtitle 
        title("Chick Growth")
        
        # add a legend 
        legend(0, 400, unique(data$Chick), cex=0.8, col=colors,
               pch=plotchar, lty=linetype, title="Chick ID")
    })
    
    
    output$stats <- renderText({
        
        data <- switch(input$DietSelect,
                       "Diet 1" = diet1,
                       "Diet 2" = diet2,
                       "Diet 3" = diet3,
                       "Diet 4" = diet4)
        
        dataT <- data %>%
            group_by(Time) %>%
            summarize(mean=mean(weight),n=n())
        
        dataT$mean[12]-dataT$mean[1]
    })
    
}

shinyApp(ui = ui, server = server)
```

